

---
title: "VCF linear model"
author: "Bastei"
date: "Friday, January 16, 2015"
output: html_document
---




# Data exploration & preprocessing




Loading required library
```{r}
library(raster)
```

Loading Landsat files:
```{r}
load("data/GewataB1.rda")
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")

Gewata <- stack(GewataB1,GewataB2,GewataB3,GewataB4,GewataB5,GewataB7, vcfGewata)
names(Gewata) <- c('band1', 'band2', 'band3', 'band4', 'band5', 'band7', 'VCF')
``` 

Histograms of Gewata show that in many bands outliers are present:
```{r, echo=FALSE}
hist(Gewata)
```

Clearing plot area
```{r}
dev.off()
```

Therefore, all values higher or lower than the mean +/- 5*sd are set to NA:
```{r}
mean <- cellStats(x = Gewata, stat = 'mean')
sd5 <- 5*(cellStats(x = Gewata, stat = 'sd'))
lowerbound <- mean - sd5
upperbound <- mean + sd5
Gewata[Gewata > upperbound] <- NA
Gewata[Gewata < lowerbound] <- NA
```

The new histograms look much better:

```{r, echo=FALSE}
hist(Gewata)
```

Clearing plot area
```{r}
dev.off()
```

Plots that demonstrate the relationship between the Landsat bands and the VCF tree cover:

```{r,echo=FALSE}
plot(Gewata[[1:6]], Gewata[[7]], maxpixels = 10000)
```

*From the plots we conclude that band 4 does not seem to explain much of the variance in VCF, while the other bands seem to have a negative relationship with VCF.*


We create a dataframe without rows containing NA values
```{r}
GewataDF <- as.data.frame(Gewata)
GewataDF <- na.omit(GewataDF)
```


# Linear Model


Creating multilinear model for VCF prediction based on bands as covariables 
```{r} 
modelLM <- lm(VCF ~ band1 + band2 + band3 + band4 + band5 + band7, data = GewataDF) 
```
model summary:
```{r,echo=FALSE}
summary(modelLM)
```
Model coefficients:
```{r,echo=FALSE}
modelLM
```


# Model quality


First, we make a new VCF raster, based on prediction modelLM:
```{r}
covs <- Gewata[[1:6]]
VCFpred <- predict(covs, model = modelLM, na.rm = T)
```

Comparison of original VCF map and predicted VCF map:

```{r,echo=FALSE}
opar <- par(mfrow=c(1, 2))
plot(Gewata[[7]], zlim = c(0,100), main = 'VCF Original')
plot(VCFpred, zlim = c(0,100), main = 'VCF Predicted')
par(opar)
```

Calculation of differences between original VCF and predicted VCF and RMSE and its plot

```{r}
VCFres <- VCFpred - Gewata[[7]]
```

```{r,echo=FALSE}
plot(VCFres, main = 'Residuals Predicted - Original')
```

**RMSE calculation**

```{r}
VCFresDF <- as.data.frame(VCFres)
VCFresDF <- na.omit(VCFresDF)
RMSE <- (mean(VCFresDF**2))**0.5
```

**Result**
```{r,echo=FALSE}
RMSE  # Tadaah!
```


# Calculation RMSE per class


Loading training data of classes:
```{r}
load("data/trainingPoly.rda")
```

Converting classes to integer values:

```{r}
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
```

Assigning 'Code' values to raster cells (where they overlap):
```{r}
classes <- rasterize(trainingPoly, Gewata, field='Code')
```

Map of training data
```{r}
# define a colour scale for the classes (as above)
# corresponding to: cropland, forest, wetland
cols <- c("orange", "dark green", "light blue")

plot(classes, col=cols, legend=FALSE)
# add a customized legend
legend("topright", legend=c("cropland", "forest", "wetland"), fill=cols, bg="white")
```

**RMSE calculation per class** 

```{r}
RS_VCFres <- (VCFres**2)
RMSEclass <- zonal(RS_VCFres, classes, fun='mean', digits=0, na.rm=TRUE)
RMSEclass[,2] <- RMSEclass[,2]**0.5
RMSE_Table <- as.data.frame(RMSEclass)
labs <- c("Cropland", "Forest", "Wetland")
RMSE_Table$zone <- labs
```

Result:
```{r, echo=FALSE}
RMSE_Table

barplot(RMSE_Table$mean, main = 'RMSEs per Class', ylim = c(0,12), names.arg = labs, col = cols, xlab = 'Land Class')
```

**CONCLUSION: RMSE indicates that modelLM predicts VCF with the highest precision/accuracy for class 'forest'. Wetland is worst predicted of the three classes based on the same model.**
